Word-Referent Identification Under Multimodal Uncertainty

### 2018-04-20 Writeup/ms.tex
modif to manuscript

  \abstract{Identifying a spoken word in a referential context requires both the
ability to process and integrate multimodal input and the ability to
reason under uncertainty. How do these tasks interact with one another?
We introduce a task that allows us to examine how adults identify words
under joint uncertainty in the auditory and visual modalities. We
propose an ideal observer model of the task which provides an account of
how auditory and visual cues are combined optimally. Model predictions
are tested in three experiments where word recognition is made under two
kinds of uncertainty: category ambiguity and/or distorting noise. In all
cases, the optimal model explains much of the variance in human mean
judgments. In particular, when the signal is not distorted with noise,
participants weight the auditory and visual cues optimally, that is,
according to the relative reliability of each modality. But when one
modality has noise added to it, human perceivers systematically prefer
the unperturbed modality to a greater extent than the optimal model
does. The study provides a formal framework which helps us understand
precisely how word form and word meaning interact in word recognition
under uncertainty. Moreover it offers a first step towards a model that
accounts for form-meaning synergy in early word learning.}
  \keywords{Language understanding; audio-visual processing; word learning; speech
perception; computational modeling. \\

@@

Language uses symbols expressed in one modality (e.g., the auditory
modality, in the case of speech) to communicate about the world, which
we perceive through many different sensory modalities. Consider hearing
someone yell \enquote{bee!} at a picnic, as a honey bee buzzes around
the food. Identifying a word involves processing the auditory
information as well as other perceptual signals (e.g., the visual image
of the bee, the sound of its wings, the sensation of the bee flying by
your arm). A word is successfully identified when information from these
modalities provide convergent evidence. However, word identification
takes place in a noisy world, and the cues received through each
modality may not provide a definitive answer. On the auditory side,
individual acoustic word tokens are almost always ambiguous with respect
to the particular sequence of phonemes they represent, which is due to
the inherent variability of how a phonetic category is realized
acoustically (Hillenbrand, Getty, Clark, \& Wheeler, 1995). And some
tokens may be distorted additionally by mispronunciation or ambient
noise. Perhaps the speaker was yelling \enquote{pea} and not
\enquote{bee}. Similarly, a sensory impression may not be enough to make
a definitive identification of a visual
category.\footnote{In the general case, language can of course be visual as well as auditory, and object identification can be done through many modalities. For simplicity, we focus on audio-visual matching here.}
Perhaps the insect was a beetle or a fly instead.

How does the listener deal with uncertainty to identify the speaker's
intended word? One rigorous way to approach this question is through
conducting an \emph{ideal observer} analysis. This research strategy
provides a characterization of the task/goal and shows what the optimal
performance should be under this characterization.\footnote{It is, thus,
  a general instance of the rational approach to cognition (Anderson,
  1990). It can also be seens as an instance of Marr's computational
  level of analysis.} When there is uncertainty in the input, the ideal
observer performs an optimal probabilistic inference. For example, in
order to recognize an ambiguous linguistic input, the model uses all
available probabilistic knowledge in order to maximize the accuracy of
this recognition. When the task is well specified, the ideal observer
model can be seen as a theoretical upper limit on performance. It is
supposed to be used, not so much as a realistic model of human
performance, as much as a baseline against which human performance can
be compared (Geisler, 2003; Rahnev \& Denison, 2018). When there is a
deviation from the ideal, it can reveal extra constraints on human
cognition, such as limitations on the working memory or attentional
resources.

The ideal observer analysis has had a tremendous impact not only on
speech related research (Clayards, Tanenhaus, Aslin, \& Jacobs, 2008;
Feldman, Griffiths, \& Morgan, 2009; Kleinschmidt \& Jaeger, 2015;
Norris \& McQueen, 2008), but also on many other disciplines in the
cognitive sciences (for reviews, see Chater \& Manning, 2006; Knill \&
Pouget, 2004; Tenenbaum, Kemp, Griffiths, \& Goodman, 2011). In
particular, using this research strategy, Clayards et al. (2008)
simulated auditory uncertainty by manipulating the probability
distribution of a cue (VOT) that differentiated similar words (e.g.,
\enquote{beach} and \enquote{peach}). They found that humans were
sensitive to these probabilistic cues and their judgments closely
reflected the optimal predictions. In another work, Feldman et al.
(2009) studied the perceptual magnet effect, which is a phenomenon that
involves reduced discriminability near prototypical sounds in the native
language (Kuhl, 1991). They showed that this effect can be explained as
the consequence of optimally solving the problem of perception under
uncertainty. Both Clayards et al. (2008) and Feldman et al. (2009)
explored optimal performance under uncertainty in the auditory modality.
There is, however, extensive evidence that information from the visual
modality, such as the speaker's facial features, also influences speech
understanding (see Campbell, 2008 for a review). Bejjanki, Clayards,
Knill, and Aslin (2011) offered a mathematical characterization of how
probabilistic cues from speech and lip movements can be optimally
combined. They showed that human performance during audio-visual
phonemic labeling was consistent (at least at the qualitative level)
with the behavior of the ideal observer.

This previous research, however, did not systematically study speech
understanding when the visual information is obtained, not through the
speaker's facial features, but through the referential context. In fact,
experimental findings showed that information about the identity of the
semantic referent can be integrated with linguistic information to
resolve lexical and syntactic ambiguities in speech (e.g., Eberhard,
Spivey-Knowlton, Sedivy, \& Tanenhaus, 1995; Spivey, Tanenhaus,
Eberhard, \& Sedivy, 2002; Tanenhaus, Spivey-Knowlton, Eberhard, \&
Sedivy, 1995). To our knowledge, however, no study offered an ideal
observer analysis of word identification in such context, that is, when
the listener has to combine cues from the sound and the referent.
Imagine, for example, that someone is uncertain whether they heard
\enquote{pea} or \enquote{bee}, does this uncertainty make them rely
more on the referent (e.g., the object being pointed at)? Vice versa, if
they are not sure if they saw a bee or a fly, does it make them rely
more on the sound? More importantly, when input in both modalities is
uncertain to varying degrees, do they weight each modality according to
its relative reliability (which is the optimal strategy), or do they
over-rely on a particular modality (which is a sub-optimal strategy)?

On the face of it, the question of combining information from the sound
and the visual referent might seem similar to that of audio-visual
speech integration. Nevertheless, there are at least two fundamental
differences between these two cases, and both can influence the way the
auditory and visual cues are combined:

\noindent First, in the case of audio-visual speech, both modalities
offer information about the same underlying speech category. They may
differ only in terms of their informational reliability. In a
referential context, however, the auditory and visual modalities are
additionally different in terms of the roles they play in the
referential process: the auditory input represents the \emph{symbol}
whereas the visual input represents the \emph{meaning}. It has been
suggested that because of its referential property, speech is a
privileged signal for humans, starting in infancy (see Vouloumanos \&
Waxman, 2014 for a review).\footnote{There is a debate as to whether
  speech is privileged for children and adults for the same reasons.
  Whereas some researchers suggest that speech is privileged for both
  children and adults because of its ability to refer (e.g., Waxman \&
  Markow, 1995), others suggest that speech might \emph{not} have a
  referential status from the start. Rather, speech might be prefered by
  children only because of a low level auditory ``overshadowing'' (e.g.,
  Sloutsky \& Napolitano, 2003).} Thus, in a referential context, it is
possible that listeners do not treat the auditory and visual modalities
as equivalent sources of information. Instead, there could be a
sub-optimal bias for the auditory modality beyond what is expected from
informational reliability alone.

\noindent Second in the case of audio-visual speech, the auditory and
visual stimuli are expected to be perceptually correlated. The
expectation for this correlation is such that when there is a mismatch
between the auditory and visual input, people still integrate them into
a unified (but illusory) percept (McGurk \& MacDonald, 1976). In the
case of referential language, however, the multimodal association is by
nature \emph{arbitrary} (Greenberg, 1957; Saussure, 1916). For instance,
there is no logical/perceptual connection between the sound
\enquote{bee} and the corresponding insect. Moreover, variation in the
way the sound \enquote{bee} is pronounced is generally not expected to
correlate perceptually with variation in the shape (or any other visual
property) in the category of bees. In sum, cue combination in the case
of arbitrary audio-visual associations (word-referent) is likely to be
less automatic, more effortful, and therefore less conducive to optimal
integration than it is in the case of perceptually correlated
associations (as in audio-visual speech perception).

In the current study, we investigate how people combine cues from the
auditory and the visual modality to recognize words in a referential
context. In particular, we study how this combination is performed under
various degrees of uncertainty in both the auditory and the visual
modality. We perform a rational analysis of the task. First we propose
an ideal observer model that performs the combination in an optimal
fashion. Second we compare the predictions of the optimal model to human
responses. Humans can deviate from the ideal for several reasons. For
instance, as mentioned above, a sub-optimality can be induced by the
suggested privileged status of speech or by the arbitrariness of the
referential association. In order to study possible patterns of
sub-optimality, we compare the optimal model (which provides a normative
benchmark) to a descriptive model (which is fit to human responses).
Comparing parameter estimates between these two formulations allows us
to quantify the degree of deviation from optimality.

@@

We construct an ideal observer model that combines probabilistic
information from the auditory and visual modalities. In contrast to the
model used in most research on multisensory integration (e.g., Ernst \&
Banks, 2002)---which typically studies continuous stimuli (e.g., size,
location)---this probabilistic information in our case cannot be
characterized with \emph{sensory noise/variability}, only. Indeed, our
task involves responses over categorical variables (phonemes and
concepts), and therefore, the optimal model in our case should take into
account, not only the noise variability around an individual perceptual
estimate, but also its \emph{categorical variability}, i.e., the
uncertainty related to whether this perceptual estimate belongs to a
given category (see also Bankieris, Bejjanki, \& Aslin, 2017; Bejjanki
et al., 2011). In what follows, we describe a model that accounts for
both type of variability. First, we describe the model in the simplified
case of categorical variability only. Second, we augment this simplified
model to account for sensory noise.

@@

\noindent This assumption simply says that, given a word-object mapping,
e.g., \(W=\)(\enquote{cat}-CAT), variation in the way \enquote{cat} is
pronounced does not correlate with changes in any visual property of the
object CAT, which is a valid
assumption.\footnote{Note that this assumptions is more adequate in the case of arbitrary associations such as ours, and less so in the case of redundant association such as audio-visual speech. In the latter, variation in the pronunciation is expected to correlate, at least to some extent, with lip movements.}

@@

So far, we only accounted for categorical variability. For instance, if
the speaker generates a target production \(a_t\) from an auditory
category \(p(a_t | A) \sim N(\mu_{A}, \sigma^2_{A})\), the ideal model
assumes that it has direct access to this production token (i.e.,
\(a=a_t\)), and that all uncertainty is about the category membership of
this token. However, we might also want to account for noise in the
brain and/or in the environment. For example, the observer might not
have access to the exact produced target, but only to the target
perturbed by noise. If we assume this noise to be normally distributed,
that is, \(p(a | a_t) \sim N(a_t, \sigma^2_{N_A})\), then integrating
over \(a_t\) leads to the following simple expression:
\[ p(a | A) \sim N(\mu_{A}, \sigma^2_{A}+\sigma^2_{N_A})\] Similarly, in
the case of sensory noise in the visual modality, we get
\[ p(a | V) \sim N(\mu_{V}, \sigma^2_{V}+\sigma^2_{N_V})\] Finally,
using exactly the same derivation as above, we end up with the following
multimodal weighting scheme in the optimal combination model (Equation
1) which takes into account both categorical and sensory variability:

@@

In this Experiment, we start with testing the predictions in the case
where uncertainty is due to categorical variability only (i.e.,
ambiguity in terms of category memebership). We do not add any external
noise to the background and we assume that internal sensory noise is
negligible compared to categorical variability
(\(\sigma^2_{A} \gg \sigma^2_{N_A}\), and
\(\sigma^2_{V} \gg \sigma^2_{N_V}\)). Thus, we use the following cue
weighting scheme:

@@

In this condition, only the descriptive model is fit to the data, using
the expression of the posterior (Equation 1). Since the values of the
means are fixed, we have 3 free parameters: the variances for the visual
and the auditory modalities, respectively, and \(b\), the response bias.
The visual and auditory baselines as well as the optimal model are not
fit to bimodal data, but their predictions are tested against these
bimodal data. All these normative models use the variances derived from
the unimodal data and the bias term derived from the bimodal data.

Although the paradigm is within-subjects, we did not have enough
statistical power to fit a different model for each individual
participant. Instead, models were constructed with data collapsed across
all participants. That being said, the distribution of responses from
individual participants will also be studied. The fit was done with a
nonlinear least squares regression using the NLS package in R (Bates \&
Watts, 1988). We computed the values of the parameters, as well as their
95\% confidence intervals, through non-parametric bootstrap (using 10000
iterations).

@@

\noindent The descriptive model explained 95\% of total variance.
However, since the descriptive model was fit to the same data, there is
a risk that this high correlation is due to overfitting. To examine this
possibility, we cross-validated the model using half the responses to
predict the other half (averaging across 1000 random partitions). The
predictive power of the model remained very high (\(r^2\)=0.93).

@@

Overall, we found that the optimal model explained much of the variance
in the mean judgments, and largely more than what can be explained with
the auditory or the visual models alone. Moreover, the high value of the
coefficient of determination in the optimal model (\(r^2\)=0.89)
suggests that the population was near-optimal. However, we see in
Figure~\ref{fig:bimodal} that the mean responses deviated systematically
from the optimal prediction in that they were slightly pulled toward
chance (i.e., the probability 0.5). This is due to the increase in the
value of the variance associated with each modality. Note however that,
despite this increase in randomness, our analysis of modality preference
showed that the relative values of these variances were not different
(Figure~\ref{fig:bias}), meaning that there was no evidence for a
modality preference. Thus, 1) There was a simultaneous increase in the
values of the auditory and visual variances in the bimodal condition
compared to the unimodal condition, meaning that the bimodal input lead
to an increase in response randomness, and 2) this increased randomness
did not affect the relative weighting of both modalities, i.e., the
population was weighting modalities according to the relative
reliability predicted by the optimal model. This situation corresponds
to the first case of sub-optimally described in
Figure~\ref{fig:subOptim}.

\begin{figure}[!h]
\includegraphics[width=\textwidth]{ms_files/figure-latex/individual-1} \caption{Histograms of the values of the visual variance relative to the auditory variance in Experiment 1. Light color represents the values derived from each individual participant, and dark color represents simulated data sampled from the descriptive model.}\label{fig:individual}
\end{figure}

As we noted earlier, the model addresses the question of optimality at
the population level. However, it is important to know how individual
responses are distributed. In fact, one could think of an extreme case
where optimality at the population level would be misleading. Imagine,
for instance, that in the bimodal condition half the participants relied
exclusively on the visual modality, whereas the other half relied
exclusively on the auditory modality. This case could still lead to an
aggregate behaviour which appears optimal, but this optimality would be
spurious.

To examine this possibility, we consider the distribution of individual
cross-modal weighting in the bimodal condition (i.e.,
\(\frac{\sigma^2_{Vb}}{\sigma^2_{Ab}}\)). Using a factor of 10 as a
cut-off, we found that 5 participants relied almost exclusively on the
visual modality, and 12 relied almost exclusively on the auditory
modality. The percentage of both cases was relatively small compared to
the total number of participants (21.25\%). An additional number of
participants (N=7) relied on both modalities, but provided noisy
responses which lead to negative variances (probably due to mistaking
\enquote{same} for \enquote{different} or vice versa). When these
outliers were removed, the distribution had a rather unimodal shape
(Figure~\ref{fig:individual}). This finding indicates that the
population's near optimality is not spurious, but based mostly on
genuine cue combination at the individual level.

As a second analysis, we asked whether the observed variance in the
individual distribution was due to mere sampling errors or whether it
corresponded to a real between-subject variability. We simulated
individual responses from the posterior distribution whose parameters
were fit to the population as a whole (i.e., the descriptive posterior).
The resulting distribution is shown in Figure~\ref{fig:individual}. For
ease of comparison, the simulated distribution was superimposed to the
real distribution. We found that the real distribution was broader
(\(sd=\) 2.24) than the simulated distribution (\(sd=\) 1.19),
indicating that there was a real between-subject variation beyond
sampling errors. This means that the participants varied in terms of how
they weighted modalities: Compared to the predictions of the global
descriptive model, some participants relied more on the auditory
modality, whereas others relied more on the visual modality.

@@

We were also interested in whether noise in the auditory modality lead
more participants to rely exclusively on the visual modality at the
individual level. Using the same cut-off as in Experiment 1 (a factor of
10), the percentage of participants who relied exclusively on either
modalities was 34.41\%, which is much higher than the percentage
obtained in Experiment 1 (21.25\%). Moreover, the subset of participants
relying exclusively on the visual modality (compared to those who relied
exclusively on the auditory modality) increased from 29.41\% in
Experiment 1 to 68.75\% in Experiment 2, indicating that noise in the
auditory modality prompted more participants to rely exclusively and
disproportionately on the visual modality.

@@

In this Experiment, we added background noise to the visual modality.
Similar to Experiment 2, we were interested to know if participants
would treat this new source of uncertainty as predicted by the optimal
model, that is, according to the following weighting scheme:
\[\beta_a \propto \frac{1}{\sigma^2_{A}}\]
\[\beta_v \propto \frac{1}{\sigma^2_{V}+\sigma^2_{N_V}}.\] The
alternative hypothesis is that, just like noise in the auditory modality
lead to a preference for the visual input in Experiment 2, noise in the
visual modality would lead to a preference for the auditory input.

@@

@@

@@

@@

@@

>>>>

  \abstract{Identifying a spoken word in a referential context requires both the
ability to process and integrate multimodal input and the ability to
reason under uncertainty. How do these tasks interact with one another?
We introduce a task that allows us to examine how adults identify words
under joint uncertainty in the auditory and visual modalities. We
propose an ideal observer model which provides an account of how
auditory and visual cues are combined optimally. Model predictions are
tested in three experiments where word recognition is made under two
kinds of uncertainty: category ambiguity and/or distorting noise. In all
cases, the optimal model explains much of the variance in human mean
judgments. In particular, when the signal is not distorted with noise,
participants weight the auditory and visual cues optimally, that is,
according to the relative reliability of each modality. But when one
modality has noise added to it, human perceivers systematically prefer
the unperturbed modality to a greater extent than the optimal model
does. The study provides a formal framework which helps us understand
precisely how word form and word meaning interact in word recognition
under uncertainty. Moreover it offers a first step towards a model that
accounts for form-meaning synergy in early word learning.}
  \keywords{Language understanding; audio-visual processing; word learning; speech
perception; computational modeling. \\

@@

Language uses symbols expressed in one modality, e.g., the auditory
modality, in the case of speech, to communicate about the world, which
we perceive through many different sensory modalities. Consider hearing
someone yell \enquote{bee!} at a picnic, as a honey bee buzzes around
the food. Identifying a word involves processing the auditory
information as well as other perceptual signals (e.g., the visual image
of the bee, the sound of its wings, the sensation of the bee flying by
your arm). A word is successfully identified when information from these
modalities provide convergent evidence. However, word identification
takes place in a noisy world, and the cues received through each
modality may not provide a definitive answer. On the auditory side,
individual acoustic word tokens are almost always ambiguous with respect
to the particular sequence of phonemes they represent, which is due to
the inherent variability of how a phonetic category is realized
acoustically (Hillenbrand, Getty, Clark, \& Wheeler, 1995). And some
tokens may be distorted additionally by mispronunciation or ambient
noise. Perhaps the speaker was yelling \enquote{pea} and not
\enquote{bee}. Similarly, a sensory impression may not be enough to make
a definitive identification of a visual
category.\footnote{In the general case, language can of course be visual as well as auditory, and object identification can be done through many modalities. For simplicity, we focus on audio-visual matching here.}
Perhaps the insect was a beetle or a fly instead. How does the listener
deal with such multimodal uncertainty to recognize the speaker's
intended word?

The task of matching the sound to the corresponding visual object has
been extensively studied in the developmental literature since it is
considered to be an crucial instance of early word learning. For
example, many studies focused on how children succeed in this task
despite high referential ambiguity (Medina, Snedeker, Trueswell, \&
Gleitman, 2011; Pinker, 1989; Smith \& Yu, 2008; Suanda, Mugwanya, \&
Namy, 2014; Vlach \& Johnson, 2013; Vouloumanos, 2008; Yurovsky \&
Frank, 2015). However, even when they know the exact meanings of the
words, listeners (both children and adults) often face the task of
recognizing which word the speaker has uttered, especially under noisy
circumstances. The purpose of the current study is to explore the
special case of word recognition under uncertainty when adult observers
have access to multimodal cues from the speech and the referent.

One rigorous way to approach this question is through conducting an
\emph{ideal observer} analysis. This research strategy provides a
characterization of the task/goal and shows what the optimal performance
should be under this characterization.\footnote{It is, thus, a general
  instance of the rational approach to cognition (Anderson, 1990). It
  can also be seen as an instance of Marr's computational level of
  analysis.} When there is uncertainty in the input, the ideal observer
performs an optimal probabilistic inference. For example, in order to
recognize an ambiguous linguistic input, the model uses all available
probabilistic knowledge in order to maximize the accuracy of this
recognition. The ideal observer model can be seen as a theoretical upper
limit on performance. It is not so much a realistic model of human
performance, as much as a baseline against which human performance can
be compared (Geisler, 2003; Rahnev \& Denison, 2018). When there is a
deviation from the ideal, it can reveal extra constraints on human
cognition, such as limitations on the working memory or attentional
resources. The ideal observer analysis has had a tremendous impact not
only on speech related research (Clayards, Tanenhaus, Aslin, \& Jacobs,
2008; Feldman, Griffiths, \& Morgan, 2009; Kleinschmidt \& Jaeger, 2015;
Norris \& McQueen, 2008), but also on many other disciplines in the
cognitive sciences (for reviews, see Chater \& Manning, 2006; Knill \&
Pouget, 2004; Tenenbaum, Kemp, Griffiths, \& Goodman, 2011)

Some of these ideal-observer-based studies are closely related to the
question we are addressing in the current work. For instance, Clayards
et al. (2008) simulated auditory uncertainty by manipulating the
probability distribution of a cue (Voice Onset Time) that differentiated
similar words (e.g., \enquote{beach} and \enquote{peach}). They found
that humans were sensitive to these probabilistic cues and their
judgments closely reflected the optimal predictions. In another work,
Feldman et al. (2009) studied the perceptual magnet effect, which is a
phenomenon that involves reduced discriminability near prototypical
sounds in the native language (Kuhl, 1991). They showed that this effect
can be explained as the consequence of optimally solving the problem of
perception under uncertainty.

Besides the acoustic cues explored in Clayards et al. (2008) and Feldman
et al. (2009), there is extensive evidence that information from the
visual modality, such as the speaker's facial features, also influences
speech understanding (see Campbell, 2008 for a review). Bejjanki,
Clayards, Knill, and Aslin (2011) offered a mathematical
characterization of how probabilistic cues from speech and lip movements
can be optimally combined. They showed that human performance during
audio-visual phonemic labeling was consistent (at least at the
qualitative level) with the behavior of the ideal observer. This
previous research, however, did not systematically study speech
understanding when the visual information is obtained, not through the
speaker's facial features---as in audio-visual speech perception, but
through the referential context. In fact, experimental findings showed
that information about the identity of the semantic referent can be
integrated with linguistic information to resolve lexical and syntactic
ambiguities in speech (e.g., Eberhard, Spivey-Knowlton, Sedivy, \&
Tanenhaus, 1995; Spivey, Tanenhaus, Eberhard, \& Sedivy, 2002;
Tanenhaus, Spivey-Knowlton, Eberhard, \& Sedivy, 1995). To our
knowledge, however, no study offered an ideal observer analysis of word
identification in such context, that is, when the listener has to
combine cues from the sound and the referent.

On the face of it, the question of combining information from the sound
and the visual referent might seem similar to that of audio-visual
speech integration. Nevertheless, there are at least two fundamental
differences between these two cases, and both can influence the way the
auditory and visual cues are combined: First, in the case of
audio-visual speech, both modalities offer information about the same
underlying speech category. They may differ only in terms of their
informational reliability. In a referential context, however, the
auditory and visual modalities play different roles in the referential
process---in addition to possible differences in informational
reliability. Indeed, the auditory input represents the \emph{symbol}
whereas the visual input represents the \emph{meaning}. It has been
suggested that because of its referential property, speech is a
privileged signal for humans, starting in infancy (see Vouloumanos \&
Waxman, 2014 for a review).\footnote{There is a debate as to whether
  speech is privileged for children and adults for the same reasons.
  Whereas some researchers suggest that speech is privileged for both
  children and adults because of its ability to refer (e.g., Waxman \&
  Markow, 1995), others suggest that speech might \emph{not} have a
  referential status from the start. Rather, speech might be preferred
  by children only because of a low level auditory ``overshadowing''
  (e.g., Sloutsky \& Napolitano, 2003).} Thus, in a referential context,
it is possible that listeners do not treat the auditory and visual
modalities as equivalent sources of information. Instead, there could be
a sub-optimal bias for the auditory modality beyond what is expected
from informational reliability alone.

Second, in the case of audio-visual speech, the auditory and visual
stimuli are expected to be perceptually correlated. The expectation for
this correlation is such that when there is a mismatch between the
auditory and visual input, people still integrate them into a unified
(but illusory) percept (McGurk \& MacDonald, 1976). In the case of
referential language, however, the multimodal association is by nature
\emph{arbitrary} (Greenberg, 1957; Saussure, 1916). For instance, there
is no logical/perceptual connection between the sound \enquote{bee} and
the corresponding insect. Moreover, variation in the way the sound
\enquote{bee} is pronounced is generally not expected to correlate
perceptually with variation in the shape (or any other visual property)
in the category of bees. In sum, cue combination in the case of
arbitrary audio-visual associations (word-referent) is likely to be less
automatic, more effortful, and therefore less conducive to optimal
integration than it is in the case of perceptually correlated
associations (as in audio-visual speech perception).

\subsection{The current study}\label{the-current-study}

We investigate how people combine cues from the auditory and the visual
modality to recognize words in a referential context. In particular, we
study how this combination is performed under various degrees of
uncertainty in both the auditory and the visual modality. Imagine, for
example, that someone is uncertain whether they heard \enquote{pea} or
\enquote{bee}, does this uncertainty make them rely more on the referent
(e.g., the object being pointed at)? Vice versa, if they are not sure if
they saw a bee or a fly, does it make them rely more on the sound? More
importantly, when input in both modalities is uncertain to varying
degrees, do they weight each modality according to its relative
reliability (which is the optimal strategy), or do they over-rely on a
particular modality (which is a sub-optimal strategy)?

We perform a rational analysis of the task. First we propose an ideal
observer model that performs the combination in an optimal fashion.
Second we compare the predictions of the optimal model to human
responses. Humans can deviate from the ideal for several reasons. For
instance, as mentioned above, a sub-optimality can be induced by the
suggested privileged status of speech or by the arbitrariness of the
referential association. In order to study possible patterns of
sub-optimality, we compare the optimal model (which provides a normative
benchmark) to a descriptive model (which is fit to human responses).
Comparing parameter estimates between these two formulations allows us
to quantify the degree of deviation from optimality.

@@

We construct an ideal observer model that combines probabilistic
information from the auditory and visual modalities. In contrast to the
model used in most research on multisensory integration (e.g., Ernst \&
Banks, 2002)---which typically studies continuous stimuli (e.g., size,
location)---the probabilistic information in our case cannot be
characterized with \emph{sensory noise}, only. Indeed, our task involves
responses over categorical variables (phonemes and concepts), and
therefore, the optimal model should take into account, not only the
noise variability around an individual perceptual estimate, but also its
\emph{categorical variability}, i.e., the uncertainty related to whether
this perceptual estimate belongs to a given category (see also
Bankieris, Bejjanki, \& Aslin, 2017; Bejjanki et al., 2011). In what
follows, we describe a model that accounts for both type of variability.
First, we describe the model in the simplified case of categorical
variability only. Second, we augment this simplified model to account
for sensory noise.

@@

\noindent This assumption says that, given a word-object mapping, e.g.,
\(W=\)(\enquote{cat}-CAT), variation in the way \enquote{cat} is
pronounced does not correlate with changes in any visual property of the
object CAT, which is a valid
assumption.\footnote{Note that this assumptions is more adequate in the case of arbitrary associations such as ours, and less so in the case of redundant association such as audio-visual speech. In the latter, variation in the pronunciation is expected to correlate, at least to some extent, with lip movements.}

@@

So far, we only accounted for categorical variability. For instance, if
the speaker generates a target production \(a_t\) from an auditory
category \(p(a_t | A) \sim N(\mu_{A}, \sigma^2_{A})\), the ideal model
assumes that it has direct access to this production token (i.e.,
\(a=a_t\)), and that all uncertainty is about the category membership of
this token. However, we might also want to account for internal noise in
the brain and/or external noise in the environment. For example, the
observer might not have access to the exact produced target, but only to
the target perturbed by noise. If we assume this noise to be normally
distributed, that is, \(p(a | a_t) \sim N(a_t, \sigma^2_{N_A})\), then
integrating over \(a_t\) leads to the following simple expression:
\[ p(a | A) \sim N(\mu_{A}, \sigma^2_{A}+\sigma^2_{N_A})\] Similarly, in
the case of sensory noise in the visual modality, we get
\[ p(a | V) \sim N(\mu_{V}, \sigma^2_{V}+\sigma^2_{N_V})\] Finally,
using exactly the same derivation as above, we end up with the following
multimodal weighting scheme in the optimal combination model (Equation
1) which takes into account both categorical and sensory variability:

@@

In this Experiment, we test the predictions of the model in the case
where uncertainty is due to categorical variability only (i.e.,
ambiguity in terms of category membership). We do not add any external
noise to the background and we assume that internal sensory noise is
negligible compared to categorical variability
(\(\sigma^2_{A} \gg \sigma^2_{N_A}\), and
\(\sigma^2_{V} \gg \sigma^2_{N_V}\)). Thus, we use the following cue
weighting scheme:

@@

In this condition, only the descriptive model is fit to the data, using
the expression of the posterior (Equation 1). Since the values of the
means are fixed, we have 3 free parameters: the variances for the visual
and the auditory modalities, respectively, and \(b\), the response bias.
The visual and auditory baselines as well as the optimal model are not
fit to the bimodal data, but their predictions are tested against these
bimodal data. All these normative models use the variances derived from
the unimodal data and the bias term derived from the fit to bimodal
data.

Although the paradigm is within-subjects, we did not have enough
statistical power to fit a different model for each individual
participant. Instead, models were constructed with data collapsed across
all participants. That being said, the distribution of responses from
individual participants will also be analyzed. The fit was done with a
nonlinear least squares regression using the NLS package in R (Bates \&
Watts, 1988). We computed the values of the parameters, as well as their
95\% confidence intervals, through non-parametric bootstrap (using 10000
iterations).

@@

\noindent The descriptive model explained 95\% of total variance.
However, since the descriptive model was fit to the same data, there is
a risk that this high correlation is due to overfitting. To examine this
possibility, we cross-validated the model using half the responses to
predict the other half (averaging across 10 random partitions). The
predictive power of the model remained very high (\(r^2\)=0.93).

@@

Overall, we found that the optimal model explained much of the variance
in the mean judgments, and largely more than what can be explained with
the auditory or the visual models alone. Moreover, the high value of the
coefficient of determination in the optimal model (\(r^2\)=0.89)
suggests that the population was near-optimal. However, we see in
Figure~\ref{fig:bimodal} that the mean responses deviated systematically
from the optimal prediction in that they were slightly pulled toward
chance (i.e., the probability 0.5). This fact is due to the increase in
the value of the variance associated with each modality. Note however
that, despite this increase in randomness, our analysis of modality
preference showed that the relative values of these variances were not
different (Figure~\ref{fig:bias}), meaning that there was no evidence
for a modality preference. Thus, 1) There was a simultaneous increase in
the values of the auditory and visual variances in the bimodal condition
compared to the unimodal condition, meaning that the bimodal input lead
to an increase in response randomness, and 2) this increased randomness
did not affect the relative weighting of both modalities, i.e., the
population was weighting modalities according to the relative
reliability predicted by the optimal model. This situation corresponds
to the first case of sub-optimally described in
Figure~\ref{fig:subOptim}.

\begin{figure}[!h]
\includegraphics[width=\textwidth]{ms_files/figure-latex/individual-1} \caption{Histograms of the individual values of the visual variance relative to the auditory variance in Experiment 1. Light color represents the values derived from each individual participant, and dark color represents simulated individual responses sampled from the descriptive model.}\label{fig:individual}
\end{figure}

As we noted earlier, the model addresses the question of optimality at
the population level. However, it is important to know how individual
responses are distributed. In fact, one could think of an extreme case
where optimality at the population level would be misleading. Imagine,
for instance, that in the bimodal condition half the participants relied
exclusively on the visual modality, whereas the other half relied
exclusively on the auditory modality. This case could still lead to an
aggregate behavior which appears optimal, but this optimality would be
spurious.

To examine this possibility, we consider the distribution of individual
cross-modal weighting in the bimodal condition (i.e.,
\(\frac{\sigma^2_{Vb}}{\sigma^2_{Ab}}\)). Using a factor of 10 as a
cut-off, we found that 5 participants relied almost exclusively on the
visual modality, and 12 relied almost exclusively on the auditory
modality. The percentage of both cases was relatively small compared to
the total number of participants (21.25\%). When these outliers were
removed, the distribution had a rather unimodal shape
(Figure~\ref{fig:individual}). This finding indicates that the
population's near optimality is not spurious, but based mostly on
genuine cue combination at the individual level.

As a second analysis, we asked whether the observed variance in the
individual distribution was due to mere sampling errors or whether it
corresponded to a real between-subject variability. We simulated
individual responses from the posterior distribution whose parameters
were fit to the population as a whole (i.e., the descriptive posterior).
The resulting distribution is shown in Figure~\ref{fig:individual}. For
ease of comparison, the simulated distribution was superimposed to the
real distribution. We found that the real distribution had a standard
deviation of \(sd=\) 2.24 which was larger than that of the simulated
distribution (\(sd=\) 1.16), indicating that there was real
between-subject variation beyond sampling errors. This finding means
that the participants varied in terms of how they weighted modalities:
Compared to the predictions of the population-level model, some
participants relied more on the auditory modality, whereas others relied
more on the visual modality.

@@

We were also interested in whether noise in the auditory modality lead
more participants to rely exclusively on the visual modality at the
individual level. Using the same cut-off as in Experiment 1 (a factor of
10), the percentage of participants who relied exclusively on either
modalities was 34.41\%, which is much higher than the percentage
obtained in Experiment 1 (21.25\%). Moreover, the subset of participants
relying exclusively on the visual modality (compared to those who relied
exclusively on the auditory modality) increased from 29.41\% in
Experiment 1 to 68.75\% in Experiment 2, indicating that noise in the
auditory modality prompted more participants to rely exclusively and
disproportionately on the visual modality (see Table
\ref{tab:exclusive}).

@@

In this Experiment, we added background noise to the visual modality.
Similar to Experiment 2, we were interested to know if participants
would treat this new source of uncertainty as predicted by the optimal
model, that is, according to the following weighting scheme:
\[\beta_a \propto \frac{1}{\sigma^2_{A}}\]
\[\beta_v \propto \frac{1}{\sigma^2_{V}+\sigma^2_{N_V}}.\] The
alternative hypothesis is that noise in the visual modality would lead
to a preference for the auditory input, just like noise in the auditory
modality lead to a preference for the visual input in Experiment 2.

@@

\vspace{1em}

\fbox{\parbox[b][][c]{14cm}{\centering All data and code for these analyses are available at\ \url{https://github.com/afourtassi/WordRec}}}
\vspace{1em}

@@

\section{Acknowledgements}\label{acknowledgements}

This work was supported by a post-doctoral grant from the Fyssen
Foundation + XXX

@@

\hypertarget{ref-medina2011}{}
Medina, T., Snedeker, J., Trueswell, J., \& Gleitman, L. (2011). How
words can and cannot be learned by observation. \emph{Proceedings of the
National Academy of Sciences}, \emph{108}(22), 9014.

@@

\hypertarget{ref-pinker1989}{}
Pinker, S. (1989). \emph{Learnability and cognition: The acquisition of
argument structure}. Cambridge, MA: MIT press.

@@

\hypertarget{ref-smith08}{}
Smith, L. B., \& Yu, C. (2008). Infants rapidly learn word-referent
mappings via cross-situational statistics. \emph{Cognition},
\emph{106}(3), 1558--1568.
